---
title: "Neotoma_bryanti - occurrence data cleaning"
author: "Ethan Aberccrombie"
date: "11/4/2021"
output: html_document
---

This script is a template to clean species occurrence data for a project aiming to understand how the climatic tendencies of small-mammal communities in the sierra Nevada are changing.

This code is borrowed and inspired from Adam Smith and the "Best Practices in Species Distribution Modeling" Workshop series.

http://www.earthskysea.org/best-practices-in-species-distribution-modeling-a-workshop-in-r/

A few resources:
-http://mazamascience.com/WorkingWithData/?p=1494
-https://www.r-bloggers.com/2019/04/zooming-in-on-maps-with-sf-and-ggplot2/

https://r-spatial.org/r/2018/10/25/ggplot2-sf.html

https://stackoverflow.com/questions/50144222/how-to-mark-points-by-whether-or-not-they-are-within-a-polygon

```{r, include=FALSE}
#load packages

library(tidyverse)
#spatial data packages

library(raster)
library(sf)

#ggplot functions
library(viridis)

require(rnaturalearth)
require(units)

require(WorldClimTiles)

#Set default map projection
default_crs = sf::st_crs(4326)

#increase memory (up to 3gb on Windows 64 bit)
memory.limit(memory.limit() * 2^30)
```



```{r, include=FALSE}
#load occurrence data
#Choose occurrence data in files.
species <- read.delim('C:/Grinnell-Mammal-Community-Shifts/Data/Species Occurrence Data/Neotoma_bryanti/occurrence.txt')

#R does not know how to interpret the first column.
#Change the first column to a character.
species$gbifID <- as.character(species$gbifID)
```

Look at data types and only include those that represent 'real' collections.

```{r, include=FALSE}
#observation types
unique(species$type)

#take a subset of the data.
  #We are assuming these 'types' of data represent real collections, collected by professionals.
  species <- subset(species,
    type=='PhysicalObject' |
    type=='specimen' |
    type=='PRESERVED_SPECIMEN' |
    type=='PreservedSpecimen' |
    type=='Occurrence' |
    type=="Collection" |
    type=='Physical Object' |
    type == 'Objeto fÃsico'
  )
```

```{r, include=FALSE}
### convert species data to a spatial object for plotting

	# remove records with no long/lat
speciesNoNas <- species %>% 
  filter(!is.na(decimalLatitude) & 
           !is.na(decimalLongitude))
	
#Create a sf object	
speciesSf <- st_as_sf(x = speciesNoNas,
                      coords = c(x = 'decimalLongitude',
                                 y = 'decimalLatitude'),
                      crs = 4326)

rm(list = c('species','speciesNoNas'))
```

```{r, include=FALSE}
### get outlines of states/counties of Canada, US, Mexico

load('C:/Grinnell-Mammal-Community-Shifts/Code/Occurence Data Cleaning/GADM Canada, USA, Mexico Level 0 WGS84.rda')
load('C:/Grinnell-Mammal-Community-Shifts/Code/Occurence Data Cleaning/GADM Canada, USA, Mexico Level 1 WGS84.rda')
load('C:/Grinnell-Mammal-Community-Shifts/Code/Occurence Data Cleaning/GADM Canada, USA, Mexico Level 2 WGS84.rda')

	# remove Alaska and Hawaii because they make plotting much slower... but you may want to include Alaska if any species reside there
	states <- nam1Sp[!(nam1Sp@data$NAME_1 %in% c('Hawaii', 'Alaska')), ]
	  states <- st_as_sf(states)
	counties <- nam2Sp[!(nam2Sp@data$NAME_1 %in% c('Hawaii', 'Alaska')), ]
	  counties <- st_as_sf(counties)
	  
rm(list = c('nam0Sp','nam1Sp','nam2Sp'))	  
```

```{r, include=FALSE}
#load rangemap
#choose appropriate rangemap for each individual species.
rangeMap <- st_read('C:/Grinnell-Mammal-Community-Shifts/Data/IUCN Species Ranges/Species Ranges_Exports_QGIS/Neotoma_bryanti/Neotoma_bryanti.shp')
```

# Mapping

Map of all Observations

```{r}
#All observations.
ggplot() +
  geom_sf(data = states) +
  geom_sf(data = speciesSf, 
             aes(color = coordinateUncertaintyInMeters)) +
  scale_colour_viridis(option = "magma")
```

Before manually removing records, let's apply a few filters.

Filter data. Remove all occurrences with uncertaintity > 1000m. Remove all occurrences collected before 1970.

We would like to remove points with no coordinate uncertainty that fall outside of the IUCN rangemap.

```{r}
# Because climate is constantly changing, we need to limit our references to those collections in recent years.
# Let's start by limiting our collection to years 1970-2000
hist(speciesSf$year,
    xlab='Collection year',
    ylab='Number of records',
    main='Collection year (all years)')
```


```{r}
#Define range buffer of 80km.
buffer_distance <- as_units(80, "km")

range_buffer <- rangeMap %>% 
  st_buffer(dist = buffer_distance)
#Considering climate changes rapidly in montane environments, I am going to remove occurrence that have uncertainty greater than 1000m.
speciesSf_filtered <- speciesSf %>% 
  mutate(within_range = lengths(st_within(x = speciesSf, 
                                          y = rangeMap)),
         within_buffer = lengths(st_within(x = speciesSf, 
                                          y = range_buffer))) %>% 
  filter(year >= 1970) %>% 
  filter(coordinateUncertaintyInMeters <= 1000 | is.na(coordinateUncertaintyInMeters)) %>% 
  filter(!(coordinateUncertaintyInMeters == "NA" &
           within_range == 0))
```

Plot occurrences after filtering.

```{r}
#map the occurrences again
#All of North America.
ggplot() +
  geom_sf(data = states) +
  geom_sf(data = speciesSf_filtered, 
             aes(color = coordinateUncertaintyInMeters)) +
  scale_colour_viridis(option = "magma")
```

Crop into the appropriate region. Add county and state borders.

```{r}
#Plot
ggplot() +
  geom_sf(data = counties,
          fill = NA) +
  geom_sf(data = states,
               fill = NA,
               color = "darkred") +
  geom_sf(data = rangeMap,
          color = 'darkgreen',
          fill = 'green',
          alpha = 0.5,
          inherit.aes = FALSE) +
  geom_sf(data = range_buffer,
          color = 'yellow',
          fill = NA,
          inherit.aes = F) +
  geom_sf(data = speciesSf_filtered, 
             aes(color = within_range),
          inherit.aes = FALSE) +
  coord_sf(xlim = c(-125, -108), 
           ylim = c(22, 40), 
           expand = TRUE)
```

Other issues.

Evaluate these accordingly per individual species.

```{r}
unique(speciesSf_filtered$issue)

#Evaluate the number of observations with 'Issue'
fi <- filter(speciesSf_filtered,
            grepl(c('Text of issue here'),
                  issue
            ))
fi$gbifID

#Notes about what was removed, and how many, here.

  #Remove these records
speciesSf_filtered <- speciesSf_filtered %>% 
  filter(!(grepl('COORDINATE_ROUNDED',
               issue)) |
           grepl('Issue 2 if relevant',
                 issue))
```

# Climate Data

```{r}
#Note: the raster function was previously used to download WorldClim data.
 
#Define path to worldclim data
worldclim_path <- "E:/Worldclim Data"

#Download Data appropriate for this species.
#Tiles (insert tiles) cover the range of this species.
range_tiles <- tile_name(as_Spatial(speciesSf_filtered),
                         name = 'worldclim')

worldclim_tiles <- tile_get(tiles = range_tiles,
                            var = "bio",
                            path = worldclim_path)

#merge tiles
climate_data <- tile_merge(worldclim_tiles)

#subset data
climate_data_selected <- subset(climate_data,
                                c('bio1',
                                  'bio12'))

#fill NA cells near coasts to account for fact that some records may not fall in a cell near a coast
	for (i in 1:nlayers(climate_data_selected)) {
		climate_data_selected[[i]] <- focal(climate_data_selected[[i]],
		                  w=matrix(1, nrow=3, ncol=3),
		                  fun=mean,
		                  na.rm=TRUE,
		                  NAonly=TRUE)
	}

#Rename file to match species
writeRaster(climate_data_selected,
            "C:/Grinnell-Mammal-Community-Shifts/Data/Species Climate Data/Neotoma_bryanti_climate",
            format = 'GTiff',
            overwrite = T)

#Add path to tif file
climate_data <- stack('C:/Grinnell-Mammal-Community-Shifts/Data/Species Climate Data/Neotoma_bryanti_climate.tif')

name <- c('temperature','precipitation')
names(climate_data) <- name 
```

Remove those occurrences outside of the 95th percentile

```{r}
#Extract values from points.
#Goal: keep all points within 95% of the climate range.
speciesSf_filtered <- raster::extract(climate_data,
                speciesSf_filtered,
                sp = TRUE)

#Check histogram contains 0 precipitation. If so, keep in analysis.
hist(speciesSf_filtered$precipitation)
#min precip = 77

speciesSf_filtered <- st_as_sf(speciesSf_filtered) %>% 
  mutate(Annual_Temp = temperature / 10,
         Annual_Precip = precipitation / 1000) %>% 
  filter(Annual_Temp <= quantile(Annual_Temp, 0.95) &
           Annual_Precip <= quantile(Annual_Precip, 0.95))
```

# Final Plot

This plot illustrates those occurrences that will be used to determine the climatic preference of the species.

```{r}
ggplot() +
  geom_sf(data = states,
               fill = NA,
               color = "darkred") +
  geom_sf(data = rangeMap,
          color = 'darkgreen',
          fill = 'green',
          alpha = 0.5,
          inherit.aes = FALSE) +
    geom_sf(data = range_buffer,
          color = 'yellow',
          fill = NA,
          inherit.aes = F) +
  geom_sf(data = speciesSf_filtered, 
             aes(color = Annual_Temp),
          inherit.aes = FALSE) +
  coord_sf(xlim = c(-125, -108), 
           ylim = c(22, 40), 
           expand = TRUE) 
```