---
title: "Occurence data cleaning template"
author: "Ethan Aberccrombie"
date: "7/29/2021"
output: html_document
---

This script is a template to clean species occurrence data for a project aiming to understand how the climatic tendencies of small-mammal communities in the sierra Nevada are changing.

This code is borrowed and inspired from Adam Smith and the "Best Practices in Species Distribution Modeling" Workshop series.

http://www.earthskysea.org/best-practices-in-species-distribution-modeling-a-workshop-in-r/

A few resources:  
-http://mazamascience.com/WorkingWithData/?p=1494  
-https://www.r-bloggers.com/2019/04/zooming-in-on-maps-with-sf-and-ggplot2/

https://r-spatial.org/r/2018/10/25/ggplot2-sf.html  
https://stackoverflow.com/questions/50144222/how-to-mark-points-by-whether-or-not-they-are-within-a-polygon

```{r, include=FALSE}
#load packages
library(tidyverse)

#spatial data packages
library(raster)
library(sf)

#ggplot functions
library(viridis)

require(rnaturalearth)
require(units)

require(WorldClimTiles)

#Set default map projection
default_crs = sf::st_crs(4326)

#increase memory (up to 3gb on Windows 64 bit)
memory.limit(memory.limit() * 2^30)
```

```{r, include=FALSE}
#load occurrence data
#Choose occurrence data in files.
species <- read.delim()

#R does not know how to interpret the first column.
#Change the first column to a character.
species$gbifID <- as.character(species$gbifID)
```

Look at data types and only include those that represent 'real' collections.

```{r, include=FALSE}
#observation types
unique(species$type)

#take a subset of the data.
  #We are assuming these 'types' of data represent real collections, collected by professionals.
  species <- subset(species,
    type=='PhysicalObject' |
    type=='specimen' |
    type=='PRESERVED_SPECIMEN' |
    type=='PreservedSpecimen' |
    type=='Occurrence' |
    type=="Collection" |
    type=='Physical Object' |
    type == 'Objeto fÃsico'
  )
```

```{r, include=FALSE}
### convert species data to a spatial object for plotting

# remove records with no long/lat
speciesNoNas <- species %>% 
  filter(!is.na(decimalLatitude) & 
           !is.na(decimalLongitude))
	
#Create a sf object	
speciesSf <- st_as_sf(x = speciesNoNas,
                      coords = c(x = 'decimalLongitude',
                                 y = 'decimalLatitude'),
                      crs = 4326)

#Remove redundant objects.
rm(list = c('species','speciesNoNas'))
```

```{r, include=FALSE}
### get outlines of states/counties of Canada, US, Mexico

load('C:/Grinnell-Mammal-Community-Shifts/Code/Occurence Data Cleaning/GADM Canada, USA, Mexico Level 0 WGS84.rda')
load('C:/Grinnell-Mammal-Community-Shifts/Code/Occurence Data Cleaning/GADM Canada, USA, Mexico Level 1 WGS84.rda')
load('C:/Grinnell-Mammal-Community-Shifts/Code/Occurence Data Cleaning/GADM Canada, USA, Mexico Level 2 WGS84.rda')

# remove Alaska and Hawaii because they make plotting much slower... but you may want to include Alaska if any species reside there
	states <- nam1Sp[!(nam1Sp@data$NAME_1 %in% c('Hawaii', 'Alaska')), ]
	  states <- st_as_sf(states)
	counties <- nam2Sp[!(nam2Sp@data$NAME_1 %in% c('Hawaii', 'Alaska')), ]
	  counties <- st_as_sf(counties)
	  
rm(list = c('nam0Sp','nam1Sp','nam2Sp'))	  
```

```{r, include=FALSE}
#load rangemap
#choose appropriate rangemap for each individual species.
rangeMap <- st_read()
```

# Filter Data

-Remove records with no latitude or longitude.  
-Remove all occurrences with uncertainty > 1000m.  
-Remove points with no coordinate undertainty that fall outside of the IUCN rangemap.
-Remove all occurrences collected before 1970.  
-Remove data with specific issues in the **issues** field.

```{r}
#Define range buffer of 80km.
buffer_distance <- as_units(80, "km")

#Create range buffer polygon.
range_buffer <- rangeMap %>% 
  st_buffer(dist = buffer_distance)

#Filter data.
speciesSf_filtered <- speciesSf %>% 
  mutate(within_range = lengths(st_within(x = speciesSf, 
                                          y = rangeMap)),
         within_buffer = lengths(st_within(x = speciesSf, 
                                          y = range_buffer))) %>% 
  filter(year >= 1970) %>% 
  filter(coordinateUncertaintyInMeters <= 1000 | is.na(coordinateUncertaintyInMeters)) %>% 
  filter(!(coordinateUncertaintyInMeters == "NA" &
           within_range == 0)) %>%  
  filter(!(grepl('COORDINATE_ROUNDED',
               issue)) &
           !(grepl('RECORDED_DATE_INVALID',
                 issue)) &
           !(grepl('COORDINATE_UNCERTAINTY_METERS_INVALID',
                 issue)) &
           !(grepl('PRESUMED_NEGATED_LONGITUDE',
                 issue))
         )

#Evaluate remaining issues
unique(speciesSf_filtered$issue)
```

Plot occurrences after filtering.

```{r}
#map the occurrences again
#All of North America.
ggplot() +
  geom_sf(data = states) +
  geom_sf(data = speciesSf_filtered, 
             aes(color = coordinateUncertaintyInMeters)) +
  scale_colour_viridis(option = "magma")
```

Crop into the appropriate region. This plot illustrates those occurrences that will paired with climate data to calculate species-specific climate affinities.

```{r}
#Plot
ggplot() +
  geom_sf(data = counties,
          fill = NA) +
  geom_sf(data = states,
               fill = NA,
               color = "darkred") +
  geom_sf(data = rangeMap,
          color = 'darkgreen',
          fill = 'green',
          alpha = 0.5,
          inherit.aes = FALSE) +
  geom_sf(data = range_buffer,
          color = 'yellow',
          fill = NA,
          inherit.aes = F) +
  geom_sf(data = speciesSf_filtered, 
             aes(color = within_range),
          inherit.aes = FALSE) +
  coord_sf(xlim = c(-125, -100), 
           ylim = c(32, 55), 
           expand = TRUE)
```

# Save data

```{r}
file_name <- 'E:/Data/Species Occurrence Data/Species Occurrence Data Cleaned/species_name_occurrence_cleaned'
write_delim(speciesSf_filtered,
            file = file_name)
```